Accurate binding of sodium and calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-11-16 Manuscript/manuscript.tex

\begin{abstract}
  % insert abstract here
Binding affinities and stoichiometries of Na$^+$ and Ca$^{2+}$ ions to phospholipid
bilayers are of paramount significance in the properties and functionality of cellular
membranes. For example, possible accumulation of ions at the membrane interface determines
the interaction affinity with charged species. Current estimates of binding affinities and
stoichiometries of cations are, however, controversial due to limitations in the available experimental and
theoretical methods. Experimentally one can assess these parameters by titrating
several membrane properties as a function of the ion concentration. However, the
interpretation of the experiments relies on theoretical models, which give
the best idealization of the ion binding as the direct experiments are not available. 
Classical molecular dynamics (MD) simulations provide details of the ion binding process with atomistic
resolution, therefore offering all the necessary information to interpret
experimental data without the need to resort to simplified models. However, the accuracy
of the available lipid models when interacting with ions is not sufficient for
such interpretation. In this work, we improve the binding details of Na$^+$ and Ca$^{2+}$ ions to the
1-Palmitoyl-2-oleoyl-phosphatidylcholine (POPC) bilayer by implicitly including electronic
polarization as a mean field correction, known as the electronic continuum correction (ECC).
This is applied to the partial charges of a selected state of the art POPC lipid model for MD simulations.
Our improved ECC-POPC model reproduces not only the
experimentally measured structural parameters for the ion-free membrane, but also the
response of lipid headgroup to a bound positive charge, and the binding affinities of 
Na$^+$ and Ca$^{2+}$ ions. With our new model we also observe negligible binding of Na$^+$ ions to POPC
bilayer and stronger interactions of Ca$^{2+}$ primarily with phosphate oxygens,
in agreement with the previous interpretations of the experimental spectroscopic data.
On the other hand, the new model results in Ca$^{2+}$ ions forming complexes with one to three POPC molecules with almost
equal probabilities, suggesting more complex binding stoichiometry than simple models
previously used to interpret the NMR data.
The results of this work pave the way to quantitative MD simulations of complex
biochemical systems with realistic electrostatic interactions in the vicinity of cellular
membranes.
\end{abstract}

@@

% Here I write pseudo-article statements that will make the main argument.
% Beautiful polished sentences will be formed only afte we agree on these basic things
\section{Introduction}
Ion interactions with cellular membranes play a key role in critical biological processes~\cite{seelig90,cevc90}. Ions, especially multivalent cations, modify general properties of the membrane which ultimately modulate their embedded transmembrane proteins~\cite{cevc90,tocanne90, pabst07}. Ions also have a more direct effect by modulating lipid--, protein--, and sugar--lipid interactions at the membrane surface. For example, Ca$^{2+}$  is crucial in neural signal propagation. It promotes membrane fusion by bridging lipids in the vesicles carrying neurotransmitters and the neuron synapsis membrane~\cite{REF}. Ca$^{2+}$ also participates in the T-cell receptor activation. It induces the detachment of the positively charged cytosolic tails of the CD3 protein complex from the negatively charged intracellular membrane making it accessible for Lck protein~\cite{Shi2013}. A final example involving sugar-lipid interactions mediated by ions is that Ca$^{2+}$ can modulate the presentation of the sugars found in the PI(4,5)P$_2$ lipid which ultimately modulates phospholipase C delta 1 pleckstrin homology domain (PLC $\delta$1-H)~\cite{Bilkova2017Calcium}. Interestingly, this modulation is lost when using Mg$^{2+}$ instead of Ca$^{2+}$ illustrating the selectivity of these processes towards particular ions. Despite our increasing understanding of the role of ions in cell membrane-related processes the exact molecular details of the mechanisms behind such processes remain elusive. \todo{Hector, please deliver the missing citations for this paragraph.}

@@

According to ECC, the electronic polarizability can be modeled in non-polarizable MD by effectively embedding the ions in a homogeneous dielectric continuum with a dielectric constant $\epsilon_{el}$, which is the electronic part of the dielectric constant of the medium~\cite{leontyev11}. Following Coulomb's law,  ECC can be directly incorporated by scaling the charges with a constant scaling factor of $f_q = \epsilon _{el} ^{-1/2}$, giving
\begin{equation}
Q^{ECC} = f_q \cdot Q
\end{equation}
for the ECC corrected charges in simulation. 
Given that the  high frequency dielectric constant of water is $\epsilon _{el} = 1.78$ (\textit{i.e.,} the square of the refraction index), the scaling factor for ions in water is roughly $f_q \approx 0.75$. This scaling factor successfully improves the accuracy of simulations of solvated ions, when quantitatively compared with neutron scattering data \cite{kohagen14,kohagen16, Pluharova2014, martinek17}.
It is important to note that the 
high frequency dielectric constant 
is around 2 for almost any biologically relevant environment \cite{leontyev11}.
The dielectric discontinuity in a lipid bilayer arises
from the orientational polarization of the molecules. 
Thus, the same correction for the electronic polarizability can be 
applied throughout the lipid bilayer interface.

@@

To reproduce the experimental ion binding affinities,
we scanned the optimal interval for the scaling factor, $f_q~\in~(0.75, 1.0)$.
The ion binding affinity was evaluated against experiments
using the head group order parameters and the electrometer concept \cite{seelig87,catte16},
as discussed more detail in the next section.
In general, scaling down the partial charges reduced the ion binding affinity.
We found an optimal binding affinity with the scaling factor of $f_q = 0.8$,
which is only slightly higher than the theoretical one ($f_q=0.75$).
Directly applying
the 0.8 scaling to the partial charges of the head group, the glycerol backbone, and
the carbonyls reduced the area per lipid to 60~\AA$^2$. This area is smaller than in the
original Lipid14 model ($65.6 \pm 0.5$~\AA$^2$)~\cite{dickson14} and in experiments
(64.3~\AA$^2$) \cite{kucerka11}. The decrease of the area per lipid arises from a
reduced hydration of the lipid head group region after scaling charges which effectively
reduces head group polarity. We solve this difficulty by reducing the effective radii of
the modified atoms by lowering the $\sigma$ parameters in the Lennard-Jones potential by a
factor of $f_\sigma = 0.89$. The same solution was taken for the ECC-ions in aqueous
solvent \cite{kohagen14, kohagen16, Pluharova2014, martinek17}. After reducing the $\sigma$ parameters,
the area per molecule is restored to the experimental value (Table~\ref{tab:apls}).

@@

The relation between the amount of the bound charge per lipid, $X^\pm$, and the head group order parameter change, $\Delta S_{\rm{CH}}^{i}$, is empirically quantified as~\cite{seelig87,ferreira16}
\begin{equation}\label{OPchangeEQ}
\Delta S_{\rm{CH}}^{i}= S_{\rm{CH}}^{i}(X^\pm)-S_{\rm{CH}}^{i}(0) \approx m_i \frac{4 }{3\chi}X^\pm,
\end{equation}
where $i$ refers to either $\alpha$ or $\beta$ carbon, $S_{\rm{CH}}^{i}(0)$ denote the order parameter in the absence of bound charge, $\chi$ is the quadrupole coupling constant ($\chi \approx$\,167\,kHz), and $m_i$ is a empirical constant depending on the valency and position of the bound charge.

@@

\subsection{Salt concentrations and binding affinity}
NMR experiments report the used salt concentrations in two different ways when measuring head group order parameters.
Some use the salt concentrations in water before solvating the lipids~\cite{akutsu81}, $C_w$,
while others use atomic absorption spectroscopy and report the salt concentration in
the supernatant after the solvation of lipids~\cite{altenbach84}, $C_b$. In this work,
%we use the latter definition. In particular,
we measure the latter concentration in the aqueous
bulk region using the farthest point from the lipid bilayer in the aqueous phase.
Notice that former concentration was used by Catte et al.~\cite{catte16}.
Although these two definitions vary when applied
to Ca$^{2+}$, they do not significantly affect the conclusions of this work.

@@

\subsubsection{Simulations of POPC bilayers with aqueous ions}
Simulations of a POPC bilayer in pure water or at varying salt concentrations consisted of 128 POPC molecules
and approximately 50~water molecules per lipid in an orthorhombic simulation box with periodic boundary conditions.
The SPC/E~\cite{Berendsen1987} water model was used in all ECC-lipid model simulations reported
in the main text. This model was also used in the previous parametrization
of ECC-ions \cite{martinek17, kohagen16, Pluharova2014} and its lowered dielectric constant is consistent with
the ECC concept~\cite{leontyev11, leontyev14}.
The robustness of our approach with other water models (OPC \cite{Izadi14},
OPC3~\cite{Izadi16}, TIP3P~\cite{jorgensen83}, TIP3p-FB and TIP4p-FB~\cite{Wang2014}, and TIP4p/2005 \cite{Abascal2005})
is tested in SI.
Sodium, calcium, and chloride ions were modeled as ECC-ions
with parameters from Refs. \citenum{martinek17, Pluharova2014}. Scaled charges and Lennart-Jones radii for atoms
forming the POPC lipid are derived in this work starting from the Lipid14 force field~\cite{dickson14}.
For comparison, simulations with the Lipid14 model~\cite{dickson14} and
ion models by Dang and coworkers~\cite{smith94,chang1999,dang2006} or ECC-ions \cite{martinek17, kohagen16, Pluharova2014}
were also performed. The TIP3p water model~\cite{jorgensen83} was used in all simulations with the original Lipid14 model.
\todo{Joe, please confirm that this is correct.}
Simulation data for the Lipid14 model with \AA{}qvist ions~\cite{aqvist90} and TIP3P~\cite{jorgensen83} water models were taken directly from
Refs.~\citenum{catte16,lipid14POPC0mMNaClfiles,lipid14POPC350mMCaClfiles,lipid14POPC350mMCaClfilesNC,lipid14POPC1000mMNaClfiles}.

@@

\begin{table}[tb!]
  \caption{Values of the area per lipid (APL) of POPC bilayers without ions from the Lipid14 simulation
    ran in this work and from the literature \cite{dickson14}, from the ECC-POPC model, and from experiments. \label{tab:apls}
  }
    %    { \color{red} HECTOR:Are we using here the same water model? If not is it fair the comparison considering the large changes observed in the SI table?
    %      SAMULI: For Lipid14 simulations the water is TIP3P and SPC/E for ECC-lipids. I do understand why the comparison would
    %    not be fair in this context?}
  \begin{tabular}{l|c c}
    model          & APL (\AA$^2$)   & Temperature [K] \\
    \hline
    Lipid14                   & 65.1$\pm$ 0.6  &  300 \\
    Lipid14 \cite{dickson14}  & 65.6$\pm$ 0.5  &  303 \\
    \hline
    ECC-POPC                & 63.2$\pm$ 0.6  &  300       \\
    \hline
    experiment \cite{kucerka11} & 64.3  &  303    \\
    \hline
  \end{tabular}
\end{table}

@@

\subsection{Validation ECC-POPC model using binding affinities to Na$^+$ and Ca$^{2+}$ Cations: the electrometer concept}

@@

\begin{figure}[htb!]
  \centering
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_CaCl.pdf}
  \caption{\label{fig:delta_ordPar_CaCl}
    Changes of the head group order parameters and P-N vector orientation of a POPC bilayer 
    as a function of the CaCl$_2$ concentration in bulk ($C_b$)
    from simulations at 313 K together with experimental data 
    (DPPC (323\,K) \cite{akutsu81} and POPC (313\,K) \cite{altenbach84}). 
    %Ion concentrations in bulk water are shown in x-axis. 
    %Bulk concentrations from simulations are calculated 
    %from the farthest point from the lipid bilayer in the aqueous phase
    The error estimate for bulk concentrations was approximately 10\,mM.
    Exception was the order of magnitude larger error in
    simulation with Lipid14 and ECC-ions due to the unconverged densities  (shown if Fig.~\ref{fig:cacl-dens}) in too 
    small simulation box. 
    Simulation data with Lipid14 and \AA{}qvist ion parameters at 298 K are taken directly from
    Refs.~\cite{lipid14POPC0mMNaClfiles,lipid14POPC350mMCaClfiles,lipid14POPC350mMCaClfilesNC}.
  }
  %\todo{Experimental values from \cite{akutsu81} to be put in the $g_3$ figure:
    %the value of $g_3$ order parameter of DPPC was -0.214 
    %in the absence of ions [average of two closely spaced splittings] and -0.211 in the presence
    %of 0.35 M CaCl$_2$ (at 59 Celcius). No effect of ions could be detected on DPPC bilayers labeled
    %at the C-2 segments of both fatty acyl chains.'' }
\end{figure}

@@

In this study, we employed the electrometer concept to demonstrate that the binding of \ce{Na^+} and \ce{Ca^2+} ions to a POPC lipid bilayer can be accurately described within a classical MD simulation force field, provided that electronic polarization is implicitly included via the electronic continuum correction~\cite{leontyev11}. While the structural details of a POPC lipid bilayer simulated with the newly developed ECC-POPC model agree with experiments with an accuracy comparable to the other state of the art lipid models, the model also reproduces the experimental lipid head group order parameter responses to a cationic surfactant, NaCl, and CaCl$_2$ concentrations. It thus represents a significant improvement over other available lipid models, which overestimate cation binding affinities~\cite{catte16}. The ECC-POPC model is built upon the Lipid14 POPC model~\cite{dickson14} by scaling the partial charges by a factor of 0.8 and reducing the Lennart-Jones radii by a factor of 0.89 for the headgroup, glycerol backbone, and carbonyl atoms.

>>>>

\begin{abstract}
  % insert abstract here
Binding affinities and stoichiometries of Na$^+$ and Ca$^{2+}$ ions to phospholipid
bilayers are of paramount significance in the properties and functionality of cellular
membranes. Current estimates of binding affinities and
stoichiometries of cations are, however, controversial due to limitations in the available experimental and
theoretical methods. Experimentally one can assess these parameters by titrating
several membrane properties as a function of the ion concentration. However, the
interpretation of the experiments relies on theoretical models, which give
the best idealization of the ion binding as direct experiments are not available. 
Classical molecular dynamics (MD) simulations provide details of the ion binding process with atomistic
resolution, therefore offering all the necessary information to interpret
experimental data without the need to resort to simplified models. However, the accuracy
of the available lipid models when interacting with ions is not sufficient for
such interpretation. In this work, we improve the binding details of Na$^+$ and Ca$^{2+}$ ions to the
1-Palmitoyl-2-oleoyl-phosphatidylcholine (POPC) bilayer by implicitly including electronic
polarization as a mean field correction, known as the electronic continuum correction (ECC).
This is applied to the partial charges of a selected state of the art POPC lipid model for MD simulations.
Our improved ECC-POPC model reproduces not only the
experimentally measured structural parameters for the ion-free membrane, but also the
response of lipid headgroup to a bound positive charge, and the binding affinities of 
Na$^+$ and Ca$^{2+}$ ions. With our new model we also observe negligible binding of Na$^+$ ions to POPC
bilayer and stronger interactions of Ca$^{2+}$ primarily with phosphate oxygens,
in agreement with the previous interpretations of the experimental spectroscopic data.
The new model results in Ca$^{2+}$ ions forming complexes with one to three POPC molecules with almost
equal probabilities, suggesting more complex binding stoichiometry than simple models
previously used to interpret the NMR data.
The results of this work pave the way to quantitative MD simulations of complex
biochemical systems with realistic electrostatic interactions in the vicinity of cellular
membranes.
\end{abstract}

@@

% Here I write pseudo-article statements that will make the main argument.
% Beautiful polished sentences will be formed only afte we agree on these basic things
\section{Introduction}
Ion interactions with cellular membranes play a key role in critical biological processes~\cite{seelig90,cevc90}. Ions, especially multivalent cations, modify general properties of the membrane which ultimately modulate their embedded transmembrane proteins~\cite{cevc90,tocanne90, pabst07}. Ions also have a more direct effect by modulating lipid--, protein--, and sugar--lipid interactions at the membrane surface. For example, Ca$^{2+}$  is crucial in neural signal propagation. It promotes membrane fusion by bridging lipids in the vesicles carrying neurotransmitters and the neuron synapsis membrane~\cite{REF}. Ca$^{2+}$ also participates in the T-cell receptor activation. It induces the detachment of the positively charged cytosolic tails of the CD3 protein complex from the negatively charged intracellular membrane making it accessible for Lck protein~\cite{Shi2013Ca2}. A final example involving sugar-lipid interactions mediated by ions is that Ca$^{2+}$ can modulate the presentation of the sugars found in the PI(4,5)P$_2$ lipid which ultimately modulates phospholipase C delta 1 pleckstrin homology domain (PLC $\delta$1-H)~\cite{Bilkova2017Calcium}. Interestingly, this modulation is lost when using Mg$^{2+}$ instead of Ca$^{2+}$ illustrating the selectivity of these processes towards particular ions. Despite our increasing understanding of the role of ions in cell membrane-related processes the exact molecular details of the mechanisms behind such processes remain elusive. \todo{Hector: I left one unknown citation for Pavel.}

@@

According to ECC mean-field correction\todo{HECTOR: I added this back because it is an important detail.}, the electronic polarizability can be modeled in non-polarizable MD by effectively embedding the ions in a homogeneous dielectric continuum with a dielectric constant $\epsilon_{el}$, which is the electronic part of the dielectric constant of the medium~\cite{leontyev11}. Following Coulomb's law,  ECC can be directly incorporated by scaling the charges with a constant scaling factor of $f_q = \epsilon _{el} ^{-1/2}$, giving
\begin{equation}
Q^{ECC} = f_q \cdot Q
\end{equation}
for the ECC corrected charges in simulation. 
Given that the  high frequency dielectric constant of water is $\epsilon _{el} = 1.78$ (\textit{i.e.,} the square of the refraction index), the scaling factor for ions in water is roughly $f_q \approx 0.75$. This scaling factor successfully improves the accuracy of simulations of solvated ions, when quantitatively compared with neutron scattering data \cite{kohagen14,kohagen16, Pluharova2014, martinek17}.
It is important to note that the 
high frequency dielectric constant 
is around 2 for almost any biologically relevant environment \cite{leontyev11}.
The dielectric discontinuity in a lipid bilayer arises
from the orientational polarization of the molecules. 
Thus, the same correction for the electronic polarizability can be 
applied throughout the lipid bilayer interface.

@@

To reproduce the experimental ion binding affinities,
we scanned the optimal interval for the scaling factor, $f_q~\in~(0.75, 1.0)$.
The ion binding affinity was evaluated against experiments
using the head group order parameters and the electrometer concept \cite{seelig87,catte16},
as discussed more detail in the next section.
Scaling down the partial charges reduced the ion binding affinity.
We found an optimal binding affinity with the scaling factor of $f_q = 0.8$,
which is only slightly higher than the theoretical one ($f_q=0.75$).
Directly applying
the 0.8 scaling to the partial charges of the head group, the glycerol backbone, and
the carbonyls reduced the area per lipid to 60~\AA$^2$. This area is smaller than in the
original Lipid14 model ($65.6 \pm 0.5$~\AA$^2$)~\cite{dickson14} and in experiments
(64.3~\AA$^2$) \cite{kucerka11}. The decrease of the area per lipid arises from a
reduced hydration of the lipid head group region after scaling charges which effectively
reduces head group polarity. We solve this difficulty by reducing the effective radii of
the modified atoms by lowering the $\sigma$ parameters in the Lennard-Jones potential by a
factor of $f_\sigma = 0.89$. The same solution was taken for the ECC-ions in aqueous
solvent \cite{kohagen14, kohagen16, Pluharova2014, martinek17}. After reducing the $\sigma$ parameters,
the area per molecule is restored to the experimental value (Table~\ref{tab:apls}).

@@

The relation between the amount of the bound charge per lipid, $X^\pm$, and the head group order parameter change, $\Delta S_{\rm{CH}}^{i}$, is empirically quantified as~\cite{seelig87,ferreira16}
\begin{equation}\label{OPchangeEQ}
\Delta S_{\rm{CH}}^{i}= S_{\rm{CH}}^{i}(X^\pm)-S_{\rm{CH}}^{i}(0) \approx m_i \frac{4 }{3\chi}X^\pm,
\end{equation}
where $i$ refers to either $\alpha$ or $\beta$ carbon, $S_{\rm{CH}}^{i}(0)$ denote the order parameter in the absence of bound charge, $\chi$ is the quadrupole coupling constant ($\chi \approx$\,167\,kHz), and $m_i$ is an empirical constant depending on the valency and position of the bound charge.

@@

\subsection{Salt concentrations and binding affinity}
NMR experiments report the used salt concentrations in two different ways when measuring head group order parameters.
Some use the salt concentrations in water before solvating the lipids~\cite{akutsu81}, $C_w$,
while others use atomic absorption spectroscopy and report the salt concentration in
the supernatant after the solvation of lipids~\cite{altenbach84}, $C_b$. In this work,
%we use the latter definition. In particular,
we measure the latter concentration in the aqueous
bulk region using the farthest point from the lipid bilayer in the aqueous phase.
Notice that the former concentration was used by Catte et al.~\cite{catte16}.
Although these two definitions vary when applied
to Ca$^{2+}$, they do not significantly affect the conclusions of this work.

@@

\subsubsection{Simulations of POPC bilayers with aqueous ions}
Simulations of a POPC bilayer in pure water or at varying salt concentrations consisted of 128 POPC molecules
and approximately 50~water molecules per lipid in an orthorhombic simulation box with periodic boundary conditions.
The SPC/E~\cite{Berendsen1987} water model was used in all ECC-POPC model simulations reported
in the main text. This model was also used in the previous parametrization
of ECC-ions \cite{martinek17, kohagen16, Pluharova2014} because its lowered dielectric constant is consistent with
the ECC concept~\cite{leontyev11, leontyev14}.
The robustness of our approach with other water models (OPC \cite{Izadi14},
OPC3~\cite{Izadi16}, TIP3P~\cite{jorgensen83}, TIP3p-FB and TIP4p-FB~\cite{Wang2014}, and TIP4p/2005 \cite{Abascal2005})
is tested in the SI.
Sodium, calcium, and chloride ions were modeled as ECC-ions
with parameters from Refs. \citenum{martinek17, Pluharova2014}. Scaled charges and Lennard-Jones radii for atoms
forming the POPC lipid are derived in this work starting from the Lipid14 force field~\cite{dickson14}.
For comparison, simulations with the Lipid14 model~\cite{dickson14} and
ion models by Dang and coworkers~\cite{smith94,chang1999,dang2006} or ECC-ions \cite{martinek17, kohagen16, Pluharova2014}
were also performed. The TIP3p water model~\cite{jorgensen83} was used in all simulations with the original Lipid14 model.
\todo{Joe, please confirm that this is correct.}
Simulation data for the Lipid14 model with \AA{}qvist ions~\cite{aqvist90} and TIP3P~\cite{jorgensen83} water models were taken directly from
Refs.~\citenum{catte16,lipid14POPC0mMNaClfiles,lipid14POPC350mMCaClfiles,lipid14POPC350mMCaClfilesNC,lipid14POPC1000mMNaClfiles}.

@@

\begin{table}[tb!]
  \caption{Values of the area per lipid (APL) of POPC bilayers without ions. \label{tab:apls}
  }
    %    { \color{red} HECTOR:Are we using here the same water model? If not is it fair the comparison considering the large changes observed in the SI table?
    %      SAMULI: For Lipid14 simulations the water is TIP3P and SPC/E for ECC-lipids. I do understand why the comparison would
    %    not be fair in this context?}
  \begin{tabular}{l|c c}
    model          & APL (\AA$^2$)   & Temperature [K] \\
    \hline
    Lipid14                   & 65.1$\pm$ 0.6  &  300 \\
    Lipid14 \cite{dickson14}  & 65.6$\pm$ 0.5  &  303 \\
    \hline
    ECC-POPC                & 63.2$\pm$ 0.6  &  300       \\
    \hline
    experiment \cite{kucerka11} & 64.3  &  303    \\
    \hline
  \end{tabular}
\end{table}

@@

\subsection{Validation ECC-POPC model using binding affinities to Na$^+$ and Ca$^{2+}$ cations: the electrometer concept}

@@

\begin{figure}[htb!]
  \centering
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_CaCl.pdf}
  \caption{\label{fig:delta_ordPar_CaCl}
    Changes of the head group order parameters and P-N vector orientation of a POPC bilayer 
    as a function of the CaCl$_2$ concentration in bulk ($C_b$)
    from simulations at 313 K together with experimental data 
    (DPPC (323\,K) \cite{akutsu81} and POPC (313\,K) \cite{altenbach84}). 
    %Ion concentrations in bulk water are shown in x-axis. 
    %Bulk concentrations from simulations are calculated 
    %from the farthest point from the lipid bilayer in the aqueous phase
    The error estimate for bulk concentrations was approximately 10\,mM.
    Exception was the order of magnitude larger error in
    simulation with Lipid14 and ECC-ions due to the unconverged densities  (shown if Fig.~\ref{fig:cacl-dens}) because the 
    small simulation box. 
    Simulation data with Lipid14 and \AA{}qvist ion parameters at 298 K are taken directly from
    Refs.~\cite{lipid14POPC0mMNaClfiles,lipid14POPC350mMCaClfiles,lipid14POPC350mMCaClfilesNC}.
  }
  %\todo{Experimental values from \cite{akutsu81} to be put in the $g_3$ figure:
    %the value of $g_3$ order parameter of DPPC was -0.214 
    %in the absence of ions [average of two closely spaced splittings] and -0.211 in the presence
    %of 0.35 M CaCl$_2$ (at 59 Celcius). No effect of ions could be detected on DPPC bilayers labeled
    %at the C-2 segments of both fatty acyl chains.'' }
\end{figure}

@@

In this study, we employed the electrometer concept to demonstrate that the binding of \ce{Na^+} and \ce{Ca^2+} ions to a POPC lipid bilayer can be accurately described within a classical MD simulation force field, provided that electronic polarization is implicitly included via the electronic continuum correction~\cite{leontyev11}. While the structural details of a POPC lipid bilayer simulated with the newly developed ECC-POPC model agree with experiments with an accuracy comparable to the other state of the art lipid models, the model also reproduces the experimental lipid head group order parameter responses to a cationic surfactant, NaCl, and CaCl$_2$ concentrations. It thus represents a significant improvement over other available lipid models, which overestimate cation binding affinities~\cite{catte16}. The ECC-POPC model is built upon the Lipid14 POPC model~\cite{dickson14} by scaling the partial charges by a factor of 0.8 and reducing the Lennard-Jones radii by a factor of 0.89 for the headgroup, glycerol backbone, and carbonyl atoms.
